Soss

### 2019-12-31 paper/paper.tex
edits

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
% \usepackage{unicode-math}
\usepackage{textgreek}
\usepackage{minted}

@@

A given inference algorithm can be expressed in terms of a collection of functions on a model. For example, we may need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples. In Soss, these \emph{inference primitives} are

@@

\section{Symbolic Simplification}
Before simplification:
\[
\begin{aligned}
    \ell = 
       &- 3 \log{\left(2 \right)} 
        - 5 \log{\left(\pi \right)} 
    \\ &- σ^{2} 
        - 4 \log{\left(λ \right)} 
        - 2 \log{\left(λ^{2} + 1 \right)} 
        - \frac{σ_I^{2}}{λ^{2}} 
        - \frac{σ_J^{2}}{λ^{2}}
    \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
    \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
    \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}}\right) 
\end{aligned}
\]

After simplification:
\[
    \begin{aligned}
    \ell =
    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J 
    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
        - σ^{2} 
        - I J \log{σ } 
    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
        - \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
    \end{aligned}
\]

>>>>

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
% \usepackage{unicode-math}
\usepackage{textgreek}
\usepackage{minted} 
\setminted{fontsize=\footnotesize}

@@

A given inference algorithm can be expressed in terms of a collection of functions on a model. For example, we may need to evaluate the log-density or its gradient, map continuous variables into $\mathbb{R}^n$, or draw random samples. In Soss, these \emph{inference primitives} generate model-specific code at execution time.

@@

\section{Symbolic Simplification}

\begin{figure}[t!]
    \centering

    \begin{subfigure}[t]{0.4\textwidth}
        \[
    \begin{aligned}
        \ell = 
           &- 3 \log{\left(2 \right)} 
            - 5 \log{\left(\pi \right)} 
            - σ^{2} 
        \\ &- 4 \log{\left(λ \right)} 
            - 2 \log{\left(λ^{2} + 1 \right)} 
            - \frac{σ_I^{2}}{λ^{2}} 
            - \frac{σ_J^{2}}{λ^{2}}
        \\ &+ \sum_{i=1}^{I} \left(- \log{\left(σ_I \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{α}_{i}^{2}}{2 σ_I^{2}}\right) 
        \\ &+ \sum_{j=1}^{J} \left(- \log{\left(σ_J \right)} - \frac{\log{\pi }}{2} - \frac{\log{2 }}{2}     - \frac{{β}_{j}^{2}}{2 σ_J^{2}}\right) 
        \\ &+ \sum_{\substack{1 \leq i \leq I \\ 1 \leq j \leq J}} \left(- \log{\left(σ \right)} - \frac    {\log{\pi }}{2} - \frac{\log{2 }}{2} - \frac{\left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}}{2 σ^{2}    }\right) 
    \end{aligned}
    \]
    \caption{Before}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.4\textwidth}
        \[
    \begin{aligned}
    \ell =
    &- 7.8 - 0.92 I - 0.92 J - 0.92 I J - σ^{2} 
    \\ &- 4 \log{λ } - 2 \log{\left(λ^{2} + 1 \right)} 
        - I J \log{σ } 
    \\ &- I \log{σ_I } - J \log{σ_J } - \frac{σ_I^{2}}{λ^{2}}  - \frac{σ_J^{2}}{λ^{2}}
    \\ &- \frac{1}{2 σ_I^{2}} \sum_{i=1}^{I} {α}_{i}^{2}
        - \frac{1}{2 σ_J^{2}} \sum_{j=1}^{J} {β}_{j}^{2}
    \\ &- \frac{1}{2 σ^{2}} \sum_{\substack{1 \leq i \leq I\\1 \leq j \leq J}} \left({y}_{ij} - {\hat{y}}_{ij}\right)^{2}
    \end{aligned}
\]
    \captionsetup{skip=27pt}
    \caption{After}
    \end{subfigure}
    \caption{Symbolic log-density of $m$ before (a) and after (b) simplification steps. Reduction }
    \label{}
\end{figure}

Before simplification:

After simplification:

\section{Extended example }
\todo[inline]{Break this up! This is just a staging area}

\begin{minted}{julia}
julia> using Revise, Soss, Random

julia> Random.seed!(1);

julia> m = @model I,J begin
           λ ~ HalfCauchy()
           σI ~ HalfNormal(λ)
           σJ ~ HalfNormal(λ)
           α ~ Normal(0, σI) |> iid(I)
           β ~ Normal(0, σJ) |> iid(J)
           yhat = α .+ β'
           σ ~ HalfNormal()
           y ~ For(I,J) do i,j 
                   Normal(yhat[i,j], σ)
               end
       end;

julia> truth = rand(m(I=2,J=3)); pairs(truth);

julia> post = dynamicHMC(m(I=2,J=3), (y=truth.y,));

julia> postpred = predict(m(I=2,J=3), post) |> particles;

julia> postpred.yhat
2×3 Array{Particles{Float64,1000},2}:
 0.241 ± 0.26  0.527 ± 0.3   -0.556 ± 0.38
 0.22 ± 0.26   0.506 ± 0.32  -0.577 ± 0.37

julia> postpred.y
2×3 Array{Particles{Float64,1000},2}:
 0.224 ± 0.58  0.529 ± 0.58  -0.58 ± 0.64
 0.223 ± 0.56  0.483 ± 0.58  -0.58 ± 0.61

julia> m.args
2-element Array{Symbol,1}:
 :I
 :J

julia> pairs(m.vals)
pairs(::NamedTuple) with 1 entry:
  :yhat => :(α .+ β')

julia> pairs(m.dists)
pairs(::NamedTuple) with 7 entries:
  :λ  => :(HalfCauchy())
  :σI => :(HalfNormal(λ))
  :σJ => :(HalfNormal(λ))
  :α  => :(Normal(0, σI) |> iid(I))
  :β  => :(Normal(0, σJ) |> iid(J))
  :σ  => :(HalfNormal())
  :y  => :(For(I, J) do i, j…

\end{minted}
